Thermodynamic structures for magic states in quantum computation

### 2021-04-15 magic_paper.tex
Corrections and tidying in appendices, left two red comments in sections that need deletion/restructuring

\begin{document}
% Distillation rate vs bound
% Proofs in App. A
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds
% Sec: Intro
% Sec: Abstract
% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?
% Eq. numbering in last line always
% British vs US english
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Magic states are necessary for achieving universal quantum computation within fault-tolerant schemes~\cite{cit:veitch, cit:mari, cit:gottesman, cit:bravyi, cit:knill, Campbell_2011}.
Identifying magic as a resource for quantum universality has led to several theories which try to provide a framework for its quantification and manipulation \cite{cit:veitch2, cit:howard, Wang_2019, Seddon_2021}.
The main question that such a theory attempts to answer is:
\begin{center}
    \emph{How do physical constraints, such as hardware limitations, noise bias or thermodynamics, restrict magic distillation protocols?}
\end{center}

@@

Given a vector $\x$ we define $\x^\downarrow$ to be the re-arrangement of the components of $\x$ into decreasing order. Now, given two $n$--component vectors $\w$ and $\r$ we define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}\label{eq:lorenz}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R:= \sum_{i=1}^n r_i$, and $\pi$ is the permutation on $n$ objects that sends $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$, where $\widetilde{w}_i := w_i/r_i$ is the vector of component-wise ratios of $\w$ and $\r$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.
We also stress that the curve contains non-differentiable points at locations where $\widetilde{\bmw}$ changes value, which we call \emph{elbows}.

@@

If we have a magic state $\rho$ that has negativity in its Wigner representation, then as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. It is easy to establish the following result.
\begin{lemma}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$, where $\rm{sn}(\rho)$ is the sum negativity of $\rho$. Thus, the majorization constraint is stronger than mana in any fragment.
\end{lemma}
\begin{proof}
	We denote the Wigner distributions of the states compactly as vectors $\bmw(\rho) \equiv W_\rho(x)$ and $\bmw(\sigma) \equiv W_\sigma(x)$.
	We choose the component indexing so that the rescaled distribution 
	\begin{equation}
		\widetilde{\bm{w}}(\rho|\sigma) \coloneqq \left(\frac{w(\rho)_1}{w(\sigma)_1}, \dots, \frac{w(\rho)_{d^2}}{w(\sigma)_{d^2}} \right)^T,
	\end{equation}
	is sorted, $\widetilde{\bm{w}} = \widetilde{\bm{w}}^\downarrow$.
	Note that all components of $\bmw(\sigma)$ are positive, so $\widetilde{w}_i \geq 0$ if and only if $w(\rho)_i \geq 0$ for any $i=1,\dots,d^2$.

@@

We now consider the task of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ as given in~\cref{eq:noisysn} into a smaller number of copies $n'$ of a less noisy Strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' < \epsilon$ and $n' \leq n$. Since we work in the unital fragment, the only free state is the maximally mixed state, and we choose to tensor in $n'-n$ auxiliary copies of it to the output. This highlights that we are working in the unital fragment, while it does not affect neither the distillation process nor our analysis, since the Lorenz curves remain unaffected. More generally, it is easy to see that
\begin{equation}
	L_{\rho |\sigma} (x) = L_{\rho \otimes \sigma |\sigma}(x),
\end{equation}
for any full-rank $\sigma \in \F$. Therefore, we consider the transformation
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
and compare the Lorenz curves for the input and output states to bound $R(\epsilon, \epsilon') \coloneqq n'/n$, the rate of this interconversion within the fragment.

To use relative majorization on such distributions we must rescale relative to the free state, however the rescaled distribution in the unital fragment is simply
\begin{equation}
	\widetilde{\rm{W}}_{\rho_{\rm{S}}|\frac{1}{3}\id}(\bmx) = \frac{\W[\bmx]{\rho_{\rm{S}}}}{\W[\bmx]{\frac{1}{3}\id}} = 9\W[\bmx]{\rho_{\rm{S}}},
\end{equation}
so the problem of ordering the rescaled distribution, which is required for the evaluation of the Lorenz curves, is equivalent to ordering $\W{\rho_{\rm{S}}}$. The Wigner distribution of $n$ copies of the state is obtained by the $n$-fold product of the single copy distribution, and the Lorenz curve $L_{\rho|\sigma}(x)$ for $\rho = \rho_S(\epsilon)^{\otimes n}$ and $\sigma = (\I/3)^{\otimes n}$ is readily computed for any $n$ and $\epsilon$, and the full details are provided in~\cref{app:lcsu_technical}.
We highlight that the non-trivial element that arises for quasi-distributions is that the product of negative components can generate a positive $n$-fold component, so the ordering of the components of $W_{\rho |\sigma}(x)$ depends on whether $-v(\epsilon)$ appears to an even or odd power in the expression below.

The component values $w_i$ and associated multiplicities ($m_i$) in the $n$--copy case are 
\begin{align}
	m_i &= 8^{i}\binom{n}{i}, \\
	w_i &= u^{i}(-v)^{n-i},
\end{align}
where index $i$ runs through $0, \dots, n$, and we assume for simplicity of exposition that $n$ is even and the noise level is not high, $\epsilon \leq 3/7$, so that $v \geq u$.
The details of this scenario, along with the other possible scenarios that occur depending on the parity of the number of copies and the noise level, are provided in~\cref{app:lcsu_coord_elb}.

@@

It is possible to derive distillation bounds from any location of the Lorenz curve, including the region around the peak, by imposing the constraint that the Lorenz curve of the input state never dips below the curve of the output.
However, it turns out that a simple and sufficiently non-trivial constraint can be obtained by considering the first elbow of the Lorenz curves. 
The largest Wigner component of occurs for $i=0$ and is unique, so the first elbow coordinates for $\rho_S(\epsilon)^{\otimes n}$ are given by $(1/9^n, v(\epsilon)^n)$, while for $\rho_S(\epsilon')^{\otimes n'}$ the first elbow occurs at $(1/9^{n'}, v(\epsilon')^{n'})$.

The first elbow constraint involves a simple interpolation, done in~\cref{app:elb_constraints}, in order to find the coordinates of the point which is vertically aligned with the first elbow of the input state, leading to the constraint
\begin{equation}
	R(\epsilon, \epsilon') \leq \frac{\ln{(3-4\epsilon)}}{\ln{(3-4\epsilon')}}.
\end{equation}

Specifically for the limiting case of distilling pure magic states ($\epsilon'=0$), we obtain an upper bound in the unital fragment given by
\begin{equation}
	R \leq 1 + \frac{\ln (1 - \frac{4}{3} \epsilon)}{\ln 3}.
\end{equation}

@@

The Strange state rescaled distribution for a magic state $\rho_S(\epsilon)$ is now given by 
\begin{equation}
	\widetilde{\rm{W}}_{\rho_{\rm{S}}|\sigma}(\bmx) = \frac{\W[\bmx]{\rho_{\rm{S}}}}{\W[\bmx]{\sigma}},
\end{equation}
which attains four distinct values on the phase space, naturally splitting it into four regions as illustrated in the~\cref{fig:pd_split_thermal}.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{figs/pd_split_thermal.pdf}
    \caption{\textbf{Qutrit phase space regions for $W_{\rho_{\rm{S}} | \sigma}(\x)$.}
    Here, the negative component of the magic state overlaps the Wigner distribution of $|0\>$. The rescaled distribution attains a single value in each of the four regions, proportional to the value depicted in the region, see~\cref{eq:bmw_rescaled}.
    }
    \label{fig:pd_split_thermal}
\end{figure}

@@

In order to construct the $n$--copy Lorenz curve $\lc{\rho_{\rm{S}}^{\otimes n}}{\sigma}$ we need to order the components of the distribution, $w(\rho_{\rm{S}} | \sigma)_{ijk}$ in decreasing order.
In particular, to find the coordinates of the first elbow $(x_0, L_0)$, we need to evaluate the maximum component,
\begin{align}
	&w(\rho_{\rm{S}} | \sigma)_{\rm{max}} \coloneqq (3\Z)^{n}\times \nonumber\\
	&\max\limits_{i,j,k}\Big\{ (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )} \Big\}, \label{eq:max_slope}
\end{align}
where $0 \leq i,j,k \leq n$ and $\alpha \coloneqq i+j+k \leq n$.
For $0 \leq \epsilon \leq 3/7$, we have $v \geq u$. Since we assume that $n$ is even, we need the sum $\alpha = i+j+k$ to be even too, so that the expression is positive.

>>>>

\begin{document}
% Write bounds as 1 + ln{}/{} or 1 - (-ln{})/{} ?
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds

% Find more (all) refs
% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?

% Eq. numbering in last line always
% UK vs US english
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Magic states are necessary for achieving universal quantum computation within fault-tolerant schemes~\cite{cit:veitch, cit:mari, cit:gottesman, cit:bravyi, cit:knill, Campbell_2011}.
Identifying magic as a resource for quantum universality has led to several theories which try to provide a framework for its quantification and manipulation \cite{cit:veitch2, cit:howard, Wang_2019, Seddon_2021, Okay_2021}.
The main question that such a theory attempts to answer is:
\begin{center}
    \emph{How do physical constraints, such as hardware limitations, noise bias or thermodynamics, restrict magic distillation protocols?}
\end{center}

@@

Given a vector $\x$ we define $\x^\downarrow$ to be the re-arrangement of the components of $\x$ into decreasing order. Now, given two $n$--component vectors $\w$ and $\r$ we define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}
\label{eq:lc}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R:= \sum_{i=1}^n r_i$, and $\pi$ is the permutation on $n$ objects that sends $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$, where $\widetilde{w}_i := w_i/r_i$ is the vector of component-wise ratios of $\w$ and $\r$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.
We also stress that the curve contains non-differentiable points at locations where $\widetilde{\bmw}$ changes value, which we call \emph{elbows}.

@@

If we have a magic state $\rho$ that has negativity in its Wigner representation, then as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. In fact, we can establish the following general result.
\begin{lemma}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$, where $\rm{sn}(\rho)$ is the sum negativity of $\rho$. Thus, the majorization constraint is stronger than mana in any fragment.
\end{lemma}
\begin{proof}
	We denote the Wigner distributions of the states compactly as vectors $\bmw(\rho) \equiv W_\rho(x)$ and $\bmw(\sigma) \equiv W_\sigma(x)$.
	We choose the component indexing so that the rescaled distribution 
	\begin{equation}
		\widetilde{\bm{w}}(\rho|\sigma) \coloneqq \left(\frac{w(\rho)_1}{w(\sigma)_1}, \dots, \frac{w(\rho)_{d^2}}{w(\sigma)_{d^2}} \right)^T,
	\end{equation}
	is sorted, $\widetilde{\bm{w}} = \widetilde{\bm{w}}^\downarrow$.
	Note that all components of $\bmw(\sigma)$ are positive, so $\widetilde{w}_i \geq 0$ if and only if $w(\rho)_i \geq 0$ for any $i=1,\dots,d^2$.

@@

We now consider the task of purifying $n$ copies of a noisy Strange state $\rho_{\rm{S}}(\epsilon)^{\otimes n}$ as given in~\cref{eq:noisysn} into a smaller number of copies $n'$ of a less noisy Strange state $\rho_{\rm{S}}(\epsilon')^{\otimes n'}$, with $\epsilon' < \epsilon$ and $n' \leq n$. Since we work in the unital fragment, the maximally mixed state is free and we choose to tensor in $n'-n$ auxiliary copies of it to the output, highlighting that we are working in the unital fragment. This affects neither the distillation process nor our analysis, since the Lorenz curves remain unaffected. More generally, it is easy to see that
\begin{equation}
	L_{\rho |\sigma} (x) = L_{\rho \otimes \sigma |\sigma \otimes \sigma}(x),
\end{equation}
for any full-rank $\sigma \in \F$. Therefore, we consider the transformation
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n-n')},
\end{equation}
and compare the Lorenz curves for the input and output states to bound $R(\epsilon, \epsilon') \coloneqq n'/n$, the rate of this interconversion within the fragment.

To use relative majorization on such distributions we must rescale relative to the free state, however the rescaled distribution in the unital fragment is simply
\begin{equation}
	\widetilde{\rm{W}}_{\rho_{\rm{S}}|\frac{1}{3}\id}(\bmx) = \frac{\W[\bmx]{\rho_{\rm{S}}}}{\W[\bmx]{\frac{1}{3}\id}} = 9\W[\bmx]{\rho_{\rm{S}}},
\end{equation}
so the problem of ordering the rescaled distribution, which is required for the evaluation of the Lorenz curves, is equivalent to ordering $\W{\rho_{\rm{S}}}$. The Wigner distribution of $n$ copies of the state is obtained by the $n$-fold product of the single copy distribution, and the Lorenz curve $L_{\rho|\sigma}(x)$ for $\rho = \rho_S(\epsilon)^{\otimes n}$ and $\sigma = (\I/3)^{\otimes n}$ is readily computed for any $n$ and $\epsilon$, and the full details are provided in~\cref{app:lcsu_technical}.
We highlight that the non-trivial element that arises for quasi-distributions is that the product of negative components can generate a positive $n$-fold component, so the ordering of the components of $W_{\rho |\sigma}(x)$ depends on whether $-v(\epsilon)$ appears to an even or odd power in the expression of~\cref{eq:wigu}.

The component values $w_i$ and associated multiplicities ($m_i$) in the $n$--copy case are 
\begin{align}
	m_i &= 8^{i}\binom{n}{i}, \\
	w_i &= u^{i}(-v)^{n-i}, \label{eq:wigu}
\end{align}
where index $i$ runs through $0, \dots, n$, and we assume for simplicity of exposition that $n$ is even and the noise level is not high, $\epsilon \leq 3/7$, so that $v \geq u$.
The details of this scenario, along with the other possible cases that occur depending on the noise level and the parity of the number of copies, are provided in~\cref{app:lcsu_technical}.

@@

It is possible to derive distillation bounds from any location of the Lorenz curve, including the region around the peak, by imposing the constraint that the Lorenz curve of the input state never dips below the curve of the output.
However, it turns out that a simple and sufficiently non-trivial constraint can be obtained by considering the first elbow of the Lorenz curves. 
The largest Wigner component occurs for $i=0$ and is unique, so the first elbow coordinates for $\rho_S(\epsilon)^{\otimes n}$ are given by $(1/9^n, v(\epsilon)^n)$, while for $\rho_S(\epsilon')^{\otimes n'}$ by $(1/9^{n'}, v(\epsilon')^{n'})$.

The first elbow constraint involves a simple interpolation, done in~\cref{app:elb_constraints}, in order to find the coordinates of the output Lorenz curve point which is vertically aligned with the first elbow of the input Lorenz curve. 
Comparing these two points leads to the bound
\begin{equation}
	R(\epsilon, \epsilon') \leq \frac{\ln{(3-4\epsilon)}}{\ln{(3-4\epsilon')}}.
\end{equation}
For the limiting case of distilling pure magic states ($\epsilon'=0$), we obtain an upper bound in the unital fragment given by
\begin{equation}
	R \leq 1 + \frac{\ln (1 - \frac{4}{3} \epsilon)}{\ln 3}.
\end{equation}

@@

The Strange state rescaled distribution for a magic state $\rho_S(\epsilon)$ is now given by 
\begin{equation}
	\widetilde{\rm{W}}_{\rho_{\rm{S}}|\sigma}(\bmx) = \frac{\W[\bmx]{\rho_{\rm{S}}}}{\W[\bmx]{\sigma}},
\end{equation}
which attains four distinct values on the phase space, naturally splitting it into four regions as illustrated in the~\cref{fig:pd_split}.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{figs/pd_split_thermal.pdf}
    \caption{\textbf{Qutrit phase space regions for $W_{\rho_{\rm{S}} | \sigma}(\x)$.}
    Here, the negative component of the magic state overlaps the Wigner distribution of $|0\>$. The rescaled distribution attains a single value in each of the four regions, proportional to the value depicted in the region, see~\cref{eq:bmw_rescaled}.
    }
    \label{fig:pd_split}
\end{figure}

@@

In order to construct the $n$--copy Lorenz curve $\lc{\rho_{\rm{S}}^{\otimes n}}{\sigma^{\otimes n}}$ we need to order the components of the distribution, $w(\rho_{\rm{S}} | \sigma)_{ijk}$ in decreasing order.
In particular, to find the coordinates of the first elbow $(x_0, L_0)$, we need to evaluate the maximum component,
\begin{align}
	&w(\rho_{\rm{S}} | \sigma)_{\rm{max}} \coloneqq (3\Z)^{n}\times \nonumber\\
	&\max\limits_{i,j,k}\Big\{ (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )} \Big\}, \label{eq:max_slope}
\end{align}
where $0 \leq i,j,k \leq n$ and $\alpha \coloneqq i+j+k \leq n$.
For $0 \leq \epsilon \leq 3/7$, we have $v \geq u$. Since we assume that $n$ is even, we need the sum $\alpha = i+j+k$ to be even too, so that the expression is positive.
